---
title: "Arenas image sorting pipeline"
author: "Oliver Tam and Julie Lowndes"
date: "10/1/2020"
output: html_document
---

## Introduction

This document is to provide a potential workflow for checking,
renaming and potentially anonymizing image files that are labelled
according to the Arenas lab convention.

For each file name, we expect 8 sections, each separated by an underscore `_`:

1. Experimental name and researcher initials
2. Experiment date and experiment number
3. Condition and replicate number
4. Immunohistochemistry date
5. Dye, antibodies or transcript
6. Image capture date
7. Microscope type
8. Lens, zoom and image number

Example:

`HESH9-CS_201001_d11-CT0.3_181002_DAPI+rbLMX1AKN-dk555_181109_CF_10Xz2-1.tif`

Where:

- Experiment name & initial: `HESH9-CS`. 
- Experiment date (and number if provided): `181002`
- Condition (and replicate if provided): `d11-CT0.3`.
- Immunohistochemistry date: `201002`
- Dye/antibodies/transcript: `DAPI+rbLMX1AKN-dk555`.
- Image capture date: `181109`
- Microscope type: `CF`
- Lens, zoom & image number: `10X-z2-3`

## Setup

Download or clone the GitHub folder: <https://github.com/olivertam/NDCN_hackathon>. Move it somewhere you do your analyses (we can discuss strategic options later).

Open the GitHub folder from RStudio File > Open Project... or in Finder/Windows Explorer by double-clicking the `NDCN_hackathon.Rproj` file.

Navigate to the "arenas-03" folder.

We will work from this `arenasImageOrganizer.Rmd` document today. It is
an R Markdown document, which combines simple text and R code. You can
use the Table of Contents at the bottom of the file to navigate. And
please add your own notes!

We will load one library, `here`, which reduces potential filepath
issues. If you have trouble with the `here` package, an option will be
provided to run the code without it.

```{r setup}
## checks if the `here` package is installed, if not, installs 
if(! requireNamespace("here",quietly=TRUE)){
  install.packages("here")
}

## load libraries
library(here) # load `here` package
```

## Pipeline for sorting Arenas lab image files

In this demo, we will use `source` to load the various functions that
we will use.

```{r load_functions}
source(here("arenas-03","arenasImageOrganizer","R","name_checker.R"))
source(here("arenas-03","arenasImageOrganizer","R","file_mover.R"))
source(here("arenas-03","arenasImageOrganizer","R","file_anonymizer.R"))
```

```{r load_functions_no_here, eval=FALSE}
source(file.path("arenas-03","arenasImageOrganizer","R","name_checker.R"))
source(file.path("arenas-03","arenasImageOrganizer","R","file_mover.R"))
source(file.path("arenas-03","arenasImageOrganizer","R","file_anonymizer.R"))
```

The ultimate goal is to take all these functions and turn them into an
R package that you could install and load like other R packages:

```{r load_package_example, eval=FALSE}
if(! requireNamespace("devtools",quietly=TRUE)){
  install.packages("devtools")
}

install_github("ndcn_compbio/arenasImageOrganizer")

library(arenasImageOrganizer)
```

In this exercise, we have our files located in the `files` subfolder
in the `arenas-03` folder, and we want to be able to check that they
are named correctly, and then move them into the `moved` subfolder,
sorted into further subfolders corresponding to the file name. We
would also try to move them into the `anonymized` subfolder, and
anonymize the file names after they are moved into their corresponding subfolders.

First, we want to set up a folder for logging our runs. The
simplest thing to do is to create one called `logs`

```{r make_log_folder}
log_path <- "arenas-03/logs"
if(! dir.exists(here(log_path))){
    dir.create(here(log_path))
}
```

```{r make_log_folder_no_here, eval=FALSE}
log_path <- "arenas-03/logs"
if(! dir.exists(file.path(log_path))){
    dir.create(file.path(log_path))
}
```

You should hopefully now see a new folder called `logs` in the
`arenas-03` folder

### Check file names using `name_checker()`

First, we want to make sure that the files that we want to move around
are named according to the convention set up by the Arenas lab. We
will be using the `name_checker()` function that was developed in
Hackathon 2. 

Just to remind you what the `name_checker()` function requires:

```{r name_checker_usage}
args(name_checker)
```

So, to check the image file names in a folder, you need to provide the
folder (in this case, called `files`) to the `name_checker()`
function, and store the output (a log of the check results):

```{r name_checker}
name_check_log  <- name_checker(here("arenas-03","files"))
```

```{r name_checker_no_here}
name_check_log  <- name_checker(file.path("arenas-03","files"))
```

Note that since the parameter `print2screen` is `TRUE` by default, you
will get an output of the checked results on your screen as well. If
you don't want it to print to the screen (especially if there are a
lot of files and it fills up your workspace), you can add
`print2screen=FALSE` as a parameter to `name_checker()`:

```{r name_checker_no_screen_output, eval=FALSE}
name_check_log  <- name_checker(here("arenas-03","files"), print2screen=FALSE)
```

It is then recommended that you save the log output from the
`name_checker()` (especially if you didn't print to screen), so you
can carefully check that your files were named correctly.

```{r write_name_checker_log}
write.table(name_check_log, here(log_path,"name_checker_log.txt"),sep="\n",quote=F,row.names=F,col.names=F)
```

```{r write_name_checker_log_no_here, eval=FALSE}
write.table(name_check_log, file.path(log_path,"name_checker_log.txt"),sep="\n",quote=F,row.names=F,col.names=F)
```

### Organizing the image files into appropriate subfolders

Once you have checked that all the files are named correctly, we can
now try to sort/organize the image files into appropriate subfolders.

Based on the Arenas lab's nomenclature, the first three sections of
the file name corresponds to the three subfolders that the file would
go into.

For example: 

`HESH9-CS_201001_d11-CT0.3_181002_DAPI+rbLMX1AKN-dk555_181109_CF_10Xz2-1.tif`

The first three sections are `HESH9-CS`, `201001`, and
`d11-CT0.3`. Thus the final location of the file would be in this
subfolder:

```
HESH9-CS/
    201001/
	    d11-CT0.3/
```

To run this, we can use a modified version of the
`batch_rename_files.R` script from Hackathon 1, now a function called
`file_mover()` (since for computer systems, moving/copying files is
functionally the same as renaming them)

Let's get an idea of what the `file_mover()` function requires:

```{r file_mover_usage}
args(file_mover)
```

You can see that there are three parameters:
1. `folder`: You need to provide the folder where the image files
   (after checking their names) are located
2. `dest_root`: This is a folder where you want to make the various
   subfolders to organize your image. By default, it is the same
   folder as you have the files, but it also allows you to separate
   the original image files from the organized files (and lets your
   check that things are copied over correctly before you delete them)
3. `anonymize`: We will go into that later.

In our example, we would take all the image files in the `files`
folder, and organize them into the corresponding subfolders within the
`moved` folder. This is the code to do this:

```{r organize_files}
file_move_log <- file_mover(here("arenas-03","files"),here("arenas-03","moved"))
write.table(file_move_log,here(log_path,"file_mover.log"),sep="\n",quote=F,row.names=F,col.names=F)
```

```{r organize_files_no_here, eval=FALSE}
file_move_log <- file_mover(file.path("arenas-03","files"),file.path("arenas-03","moved"))
write.table(file_move_log,file.path(log_path,"file_mover.log"),sep="\n",quote=F,row.names=F,col.names=F)
```

If you now look in the `moved` folder, you should now see a set of
subfolders that should correspond to the first section of the original
file name. If you traverse further inside, you should see a subfolder
for the second, and then the third, section of the file name. You
would also notice that the organized file name is now shorter, as it
no longer requires the first three sections.

A log file is also produced in the `logs` folder providing a full list
of files that were processed, and the subfolders that they were copied
to.

### Anonymizing files for blinded analyses

One idea that was suggested by the Arenas lab was the ability to
anonymize the file names so that researchers could analyze them
blindly.

This could be achieved by the `file_anonymizer()` function.

```{r file_anonymizer_usage}
args(file_anonymizer)
```

This function takes a folder (typically the folder that contains the
image files), and generate a somewhat randomized name for each of the
files within (restricted to image files). 

The current format for the "anonymized" file is the following, but
could be easily modified:

`anonymizedFile_[date it was anonymized]_["random" number]`

The "random" number is chosen by get a list of numbers corresponding
to the number of files in that folder to be anonymized, and the
randomly selecting (without replacing) the number from that list.

A log file is automatically generated in the folder (`[date of
anonymizing]_file_key.log`), which describes the original file name of
the anonymized files. By default, the original file is deleted after
it is given an anonymized name, but you can change the `deleteOrig`
value to `FALSE` to keep the file.

Here is how you can run it on a single folder:

```{r anonymize_example}
file_anon_log <- file_anonymizer(here("arenas-03","moved","HESH9-CS","201001","GFP"),deleteOrig=FALSE)
write.table(file_anon_log,here(log_path,"file_anonymize.log"),sep="\n",quote=F,row.names=F,col.names=F)
```

```{r anonymize_example_no_here, eval=FALSE}
file_anon_log <- file_anonymizer(file.path("arenas-03","moved","HESH9-CS","201001","GFP"),deleteOrig=FALSE)
write.table(file_anon_log,file.path(log_path,"file_anonymize.log"),sep="\n",quote=F,row.names=F,col.names=F)
```

If you now go into the `HESH9-CS/201001/GFP/` subfolder in `moved`,
you should see that there are two new "anonymized" files, and a log
file containing the information of their origin.

However, you might want to do the anonymization at the same time as
organizing the image files, especially if you are creating many
subfolders that you would have to process one by one with the above
method. Hence, there is an option to run `file_mover()` to also call
`file_anonymizer()`, so you can do both steps at the same time.

In this example, I will move and anonymize the files into a different
folder called `anonymized`

```{r move_and_anonymize}
file_move_anon_log <- file_mover(here("arenas-03","files"),here("arenas-03","anonymized"),anonymize=TRUE)
write.table(file_move_anon_log,here(log_path,"file_move_anonymize.log"),sep="\n",quote=F,row.names=F,col.names=F)                                
```

```{r move_and_anonymize_no_here}
file_move_anon_log <- file_mover(file.path("arenas-03","files"),file.path("arenas-03","anonymized"),anonymize=TRUE)
write.table(file_move_anon_log,file.path(log_path,"file_move_anonymize.log"),sep="\n",quote=F,row.names=F,col.names=F)                                
```

If you now navigate to the `anonymize` folder, you can see that the
subfolder structure is there, but now each file is anonymized. For
each folder, there is also a log file with the identity of the
anonymized file name.

## Future improvements

1. Package these scripts into an R package so it can be installed and
   loaded easily by users
2. Write a better R workflow/function as a one-step command to do all
   of this, and improving the logging of results (e.g. add dates to
   the log file names)
3. Adapting the file name nomenclature (especially for anonymized
   files) according to what the Arenas would like best.
4. Whatever else the Arenas lab could think of.
