---
title: "Example Workflow"
output: html_document
---

## Save images with Arenas Lab filenaming convention

At the microscope: Take pictures and save image files with naming convention

For each file name, we expect 8 sections, each separated by an underscore `_`:

1. Experimental name and researcher initials
2. Experiment date and experiment number
3. Condition and replicate number
4. Immunohistochemistry date
5. Dye, antibodies or transcript
6. Image capture date
7. Microscope type
8. Lens, zoom and image number

Example:

`HESH9-CS_201001_d11-CT0.3_181002_DAPI+rbLMX1AKN-dk555_181109_CF_10Xz2-1.tif`

Where:

- Experiment name & initial: `HESH9-CS`. 
- Experiment date (and number if provided): `181002`
- Condition (and replicate if provided): `d11-CT0.3`.
- Immunohistochemistry date: `201002`
- Dye/antibodies/transcript: `DAPI+rbLMX1AKN-dk555`.
- Image capture date: `181109`
- Microscope type: `CF`
- Lens, zoom & image number: `10X-z2-3`


## Check names, move files, optionally anonymize filenames

```{r, setup}
## checks if the `here` package is installed, if not, installs 
if(! requireNamespace("here",quietly=TRUE)){
  install.packages("here")
}

## load libraries
library(here) # load `here` package

## source Oliver's scripts
source(here("arenas-03","arenasImageOrganizer","R","name_checker.R"))
source(here("arenas-03","arenasImageOrganizer","R","file_mover.R"))
source(here("arenas-03","arenasImageOrganizer","R","file_anonymizer.R"))

```


```{r}
## check names using name_checker.R
name_check_log  <- name_checker(here("arenas-03","files"))

## move files and write log using file_mover.R
file_move_log <- file_mover(here("arenas-03","files"), here("arenas-03","moved"))
write.table(file_move_log, here(log_path,"file_mover.log"), sep="\n", quote=F, row.names=F, col.names=F)


## Optionally anonymize files using file_anonymizer.R
file_anon_log <- file_anonymizer(here("arenas-03","moved","HESH9-CS","201001","GFP"),deleteOrig=FALSE)
write.table(file_anon_log,here(log_path,"file_anonymize.log"),sep="\n",quote=F,row.names=F,col.names=F)

```




