---
title: "Arenas image sorting pipeline"
author: "Oliver Tam and Julie Lowndes"
date: "10/1/2020"
output: html_document
---

## Introduction

This document is to explain the script `file_mover.R` for moving image
files that are labelled according to the Arenas lab convention.

For each file name, we expect 8 sections, each separated by an underscore `_`:

1. Experimental name and researcher initials
2. Experiment date and experiment number
3. Condition and replicate number
4. Immunohistochemistry date
5. Dye, antibodies or transcript
6. Image capture date
7. Microscope type
8. Lens, zoom and image number

Example:

`HESH9-CS_201001_d11-CT0.3_181002_DAPI+rbLMX1AKN-dk555_181109_CF_10Xz2-1.tif`

Where:

- Experiment name & initial: `HESH9-CS`. 
- Experiment date (and number if provided): `181002`
- Condition (and replicate if provided): `d11-CT0.3`.
- Immunohistochemistry date: `201002`
- Dye/antibodies/transcript: `DAPI+rbLMX1AKN-dk555`.
- Image capture date: `181109`
- Microscope type: `CF`
- Lens, zoom & image number: `10X-z2-3`

## Setup

Download or clone the GitHub folder: <https://github.com/olivertam/NDCN_hackathon>. Move it somewhere you do your analyses (we can discuss strategic options later).

Open the GitHub folder from RStudio File > Open Project... or in Finder/Windows Explorer by double-clicking the `NDCN_hackathon.Rproj` file.

Navigate to the "arenas-03" folder.

We will work from this `file_mover_anonymizer.Rmd` document today. It is
an R Markdown document, which combines simple text and R code. You can
use the Table of Contents at the bottom of the file to navigate. And
please add your own notes!

We will load one library, `here`, which reduces potential filepath
issues. If you have trouble with the `here` package, an option will be
provided to run the code without it.

```{r setup}
## checks if the `here` package is installed, if not, installs 
if(! requireNamespace("here",quietly=TRUE)){
  install.packages("here")
}

## load libraries
library(here) # load `here` package
```

## Loading functions for sorting Arenas lab image files

In this demo, we will use `source` to load the various functions that
we will use.

```{r load_functions}
source(here("arenas-03","arenasImageOrganizer","R","file_mover.R"))
source(here("arenas-03","arenasImageOrganizer","R","file_anonymizer.R"))
```

```{r load_functions_no_here, eval=FALSE}
source(file.path("arenas-03","arenasImageOrganizer","R","file_mover.R"))
source(file.path("arenas-03","arenasImageOrganizer","R","file_anonymizer.R"))
```

In this exercise, we have our files located in the `files` subfolder
in the `arenas-03` folder. Hopefully they have all been checked with
the `name_checker()` function to ensure that they are correctly named.

The goal of the script is to move them into the `moved` subfolder (which is
the "root" of the directory tree), and then sort into further
subfolders ("branches") corresponding to the file name. 

First, we want to set up a folder for logging our commands. The
simplest thing to do is to create one called `logs`

```{r make_log_folder}
log_path <- "arenas-03/logs"
if(! dir.exists(here(log_path))){
    dir.create(here(log_path))
}
```

```{r make_log_folder_no_here, eval=FALSE}
log_path <- "arenas-03/logs"
if(! dir.exists(file.path(log_path))){
    dir.create(file.path(log_path))
}
```

You should hopefully now see a new folder called `logs` in the
`arenas-03` folder

### Organizing the image files into appropriate subfolders

Once you have checked that all the files are named correctly, we can
now try to sort/organize the image files into appropriate subfolders.

Based on the Arenas lab's nomenclature, the first three sections of
the file name corresponds to the three subfolders that the file would
go into.

For example: 

`HESH9-CS_201001_d11-CT0.3_181002_DAPI+rbLMX1AKN-dk555_181109_CF_10Xz2-1.tif`

The first three sections are `HESH9-CS`, `201001`, and
`d11-CT0.3`. Thus the final location of the file would be in this
subfolder:

```
HESH9-CS/
    201001/
	    d11-CT0.3/
```

To run this, we can use a modified version of the
`batch_rename_files.R` script from Hackathon 1, now a function called
`file_mover()` (since for computer systems, moving/copying files is
functionally the same as renaming them)

Let's get an idea of what the `file_mover()` function requires:

```{r file_mover_usage}
args(file_mover)
```

You can see that there are three parameters:
1. `folder`: You need to provide the folder where the image files
   (after checking their names) are located
2. `dest_root`: This is a folder where you want to make the various
   subfolders to organize your image. By default, it is the same
   folder as you have the files, but it also allows you to separate
   the original image files from the organized files (and lets your
   check that things are copied over correctly before you delete them)
3. `anonymize`: We will go into that later.

In our example, we would take all the image files in the `files`
folder, and organize them into the corresponding subfolders within the
`moved` folder. This is the code to do this:

```{r organize_files}
file_move_log <- file_mover(here("arenas-03", "files"), here("arenas-03","moved"))
write.table(file_move_log, here(log_path, "file_mover.log"), sep="\n", quote=F, row.names=F, col.names=F)
```

```{r organize_files_no_here, eval=FALSE}
file_move_log <- file_mover(file.path("arenas-03", "files"), file.path("arenas-03", "moved"))
write.table(file_move_log, file.path(log_path, "file_mover.log"), sep="\n", quote=F, row.names=F, col.names=F)
```

If you now look in the `moved` folder, you should now see a set of
subfolders that should correspond to the first section of the original
file name. If you traverse further inside, you should see a subfolder
for the second, and then the third, section of the file name. You
would also notice that the organized file name is now shorter, as it
no longer requires the first three sections.

A log file is also produced in the `logs` folder providing a full list
of files that were processed, and the subfolders that they were copied
to.

#### How does it work?

Let's remind ourselves what the `file_mover()` function requires:

```{r file_mover_usage}
args(file_mover)
```

You can see that there are three parameters:
1. `folder`: the folder that has your original files
2. `dest_root`: the "root" folder where you want them to go. If you don't provide it, it will be the same folder as your files
3. `anonymize`: if you want to anonymize them

We will now dive a little deeper into how it works. If you want to
follow along, the actual code is located in `arenasImageOrganizer/R`
folder

The code first checks that the `folder` (the folder with your original
files) actually exists:

```{r check_directory, eval=FALSE}
if(! dir.exists(folder)){
    log = paste(c(paste("This folder", folder, "does not exist."), "Please double-check", ""))
    return(log)
}
```

In the section after the `if` (also called the "conditional"), it runs
the `dir.exists()` function, which returns a `TRUE/FALSE` on whether
the folder exists. Then, 

```{r conditional_test}
folder = here("arenas-03","files")
folder
log = "Does the provided folder exist?"

### What is in the code
if(! dir.exists(folder)){
    log = c(log,"This folder does not exist.")
}

### What happens when we remove the '!'
if(dir.exists(folder)){
    log = c(log,"This folder actually exist.")
}
cat(log, sep="\n")
```

Next, it looks for files that match the expected Zeiss output (`*.czi`
or `*.lsm`) or TIFF (`*.tif` or `*.tiff`). If it doesn't find any
suitable files, it'll report back with an error message.

```{r find_files, eval=FALSE}
files <- setdiff(list.files(folder, pattern="\\.czi$|\\.tif$|\\.tiff$|\\.lsm$",full.names = TRUE), list.dirs(recursive = FALSE))
if(length(files) < 1){
    log = c(paste("No suitable files identified in",folder))
    return(log)
}
```

After these initial checks, it will set up a log file with the date
and time for when the script started, and then look process the list
of files. For each file, it does a quick first check to make sure the
name has 8 sections separated by an underscore `_`:

```{r quick_check_filename, eval=FALSE}
file_info <- unlist(strsplit(current_file, "_",fixed=TRUE))
if(length(file_info) != 8){
    stop("The file name does not fit the expected nomenclature. Not progressing further.")
}
```

If the name checks out, then it will identify and create the sub-directories that
the file should go to (based on the first 3 sections), and have the
new file name containing sections 4-8.

```{r setup_dir, eval=FALSE}
file_info <- unlist(strsplit(current_file, "_",fixed=TRUE))
subdir <- do.call('file.path',as.list(file_info[1:3]))
newfile <- paste(file_info[4:length(file_info)],collapse="_")

destination  <- file.path(dest_root,subdir)
if(! dir.exists(destination)){
    dir.create(destination,recursive=TRUE)
}
```

So, for example:
`HESH9-CS_201001_d11-CT0.3_201002_DAPI+rbAADC-dk488+shTH-dk555_201003_CF_20Xz1-1.tif`
will be transfered to the subfolder: `HESH9-CS/201001/d11-CT0.3`, and
named `201002_DAPI+rbAADC-dk488+shTH-dk555_201003_CF_20Xz1-1.tif`

For now, it will copy the file to the new location, but there is code
to actually move (so delete from original location). We will also
document where the file was copied to.

```{r move_copy_files, eval=FALSE}
### To move the file
file.move(files[i],file.path(destination,newfile))


### To copy the file
file.copy(files[i],file.path(destination,newfile))

destList <- c(destList,destination)
log <- c(log,paste(current_file,"moved to",subdir))
```

### Anonymizing files for blinded analyses

One idea that was suggested by the Arenas lab was the ability to
anonymize the file names so that researchers could analyze them
blindly.

This could be achieved by the `file_anonymizer()` function.

```{r file_anonymizer_usage}
args(file_anonymizer)
```

This function takes a folder (typically the folder that contains the
image files), and generate a somewhat randomized name for each of the
files within (restricted to image files). 

The current format for the "anonymized" file is the following, but
could be easily modified:

`anonymizedFile_[date it was anonymized]_["random" number]`

The "random" number is chosen by get a list of numbers corresponding
to the number of files in that folder to be anonymized, and the
randomly selecting (without replacing) the number from that list.

A log file is automatically generated in the folder (`[date of
anonymizing]_file_key.log`), which describes the original file name of
the anonymized files. By default, the original file is deleted after
it is given an anonymized name, but you can change the `deleteOrig`
value to `FALSE` to keep the file.

Here is how you can run it on a single folder:

```{r anonymize_example}
file_anon_log <- file_anonymizer(here("arenas-03","moved","HESH9-CS","201001","GFP"), deleteOrig=FALSE)
write.table(file_anon_log, here(log_path, "file_anonymize.log"), sep="\n", quote=F, row.names=F, col.names=F)
```

```{r anonymize_example_no_here, eval=FALSE}
file_anon_log <- file_anonymizer(file.path("arenas-03", "moved", "HESH9-CS", "201001", "GFP"), deleteOrig=FALSE)
write.table(file_anon_log, file.path(log_path, "file_anonymize.log"), sep="\n", quote=F, row.names=F, col.names=F)
```

If you now go into the `HESH9-CS/201001/GFP/` subfolder in `moved`,
you should see that there are two new "anonymized" files, and a log
file containing the information of their origin.

#### How does it work?

Just to remind ourselves what the `file_anonymizer()` function requires:

```{r file_anonymizer_usage}
args(file_anonymizer)
```

1. folder: the folder with the files to be anonymized
2. deleteOrig (TRUE/FALSE): do we delete the original file after
   anonymizing
   
First, the function checks that the folder you provided exists, in the
same way as the `file_mover()` function. It will then start a log
file with the date and time of the anonymization. It will then search
for all the relevant files (e.g. `*.tiff`) to be anonymize, again,
exactly the same way that `file_mover()` works.

The next step is an attempt to "randomize" the file name. My approach
is to generate numbers from 1 to N, where N is the total number of
files:

```{r counter, eval=FALSE}
counter <- 1:length(files)
```

Then, for each file, I then
"randomly" choose a number from the counter list (without replacement)
Let's actually watch this in action:

```{r sample_without_replacement}
counter <- 1:10 # Dummy example where we have 10 files
header = paste0("OrigFile",counter) # Just to remind ourselves what the original file order was
cat("This is the list of numbers\n")
counter

result <- vector()
  
cat("\nLet's run this\n")
while(length(counter) > 0){
  if(length(counter) == 1){
    ### Only 1 number left
    filenum <- counter
    result <- c(result,filenum)
    cat(paste("\nThis is the last number,",filenum),"\n")
    counter <- counter[counter != filenum]
  }else{
    ### sample "randomly" from the numbers
    filenum <- sample(counter,1)
    cat(paste("\nThis number was chosen,",filenum),"\n")
    result <- c(result,filenum)
    ### remove the chosen number
    counter <- counter[counter != filenum]
    cat("This is what is left:\n")
    print(counter)
  }
}
names(result) = header
cat("\nHere is the anonymized number list:\n")
result
```

The code above is a very long-winded way to do it, but allows me to watch, step-by-step, how the numbers are being removed. This is actually the code that was in the first version of the `file_anonymizer()` function (back in October). If you are interested to see this version, you should be able to check it out using Github.

However, there is a more concise way to do this sampling:

```{r simplified_sample_without_replacement}
counter <- 1:10
header <- paste0("OrigFile",counter)
result <- sample(counter,length(counter))
names(result) = header
result
```

The important question: Is it different everytime you run it? You don't want to "anonymize" the same way each time.
Let's test it out:

```{r multiple_sample_without_replacement}
counter <- 1:10
header = paste0("OrigFile",counter)

result <- vector()
for(i in 1:5){
  result <- sample(counter,length(counter))
  names(result) <- header
  cat(paste("Here is the anonymized number list for run", i),"\n")
  print(result)
}
```

After it gets the "random number" then generates a "anonymized" filename starting with
`anonymizedFile`, current date, the  number (all separated by
underscores `_`), and then the file extension (with the dot before
it). It will then make a copy of the file with anonymized name, and
 if you asked for the original to be deleted, delete the original
 file. It would also store the new and original name in `dict`, which
 will be used to write into a log (in the folder with the anonymized
 file) so that there is a way to re-identify the anonymized files

```{r create_anon_name, eval=FALSE}
anonDate <- format(today, format="%y%m%d")
newfile <- paste0("anonymizedFile_",anonDate,"_",filenum[i],".", extension)
output <- paste0(folder,"/", newfile)
file.copy(files[i], output)
if(deleteOrig){
    file.remove(files[i])
}
dict <- rbind(dict,c(newfile,oldfile))
```


### Move & Anonymize

Sometimes, you might want to do the anonymization at the same time as
organizing the image files, especially if you are creating many
subfolders that you would have to process one by one with the above
method. Hence, there is an option to run `file_mover()` to also call
`file_anonymizer()`, so you can do both steps at the same time.

In this example, I will move and anonymize the files into a different
folder called `anonymized`

```{r move_and_anonymize}
file_move_anon_log <- file_mover(here("arenas-03","files"), here("arenas-03", "anonymized"), anonymize=TRUE)
write.table(file_move_anon_log, here(log_path, "file_move_anonymize.log"), sep="\n", quote=F, row.names=F, col.names=F)                                
```

```{r move_and_anonymize_no_here}
file_move_anon_log <- file_mover(file.path("arenas-03", "files"), file.path("arenas-03", "anonymized"), anonymize=TRUE)
write.table(file_move_anon_log, file.path(log_path, "file_move_anonymize.log"), sep="\n", quote=F, row.names=F, col.names=F)                                
```

If you now navigate to the `anonymize` folder, you can see that the
subfolder structure is there, but now each file is anonymized. For
each folder, there is also a log file with the identity of the
anonymized file name.

#### How does it work?

This is using the third parameter from the `file_mover()` function:

```{r file_mover_usage}
args(file_mover)
```

When we set `anonymize=TRUE`, then it will run the following code in the `file_mover()` script:

```{r file_move_anon, eval=FALSE}
# We have been storing the folder names where files have been moved as part of file_mover()
destList <- c(destList, destination)

# The following will run if anonymize = TRUE
if(anonymize){
    log = c(log,"")
    # Get a list of unique folders that had a file moved to it
    destList <- unique(destList)
    
    for(i in 1:length(destList)){
      # Anonymize all files in each of the folder
      anonlog <- file_anonymizer(destList[i])
      log <- c(log,anonlog)
    }
}
```

Note that the `file_mover()` script is calling the `file_anonymizer()` script. By loading all the Arenas lab scripts, it now treats `file_anonymizer()` like other built-in R functions (e.g. `unique`).

## Future improvements

1. Package these scripts into an R package so it can be installed and
   loaded easily by users
2. Write a better R workflow/function as a one-step command to do all
   of this, and improving the logging of results (e.g. add dates to
   the log file names)
3. Adapting the file name nomenclature (especially for anonymized
   files) according to what the Arenas would like best.
4. Handle case issues with file extension
5. Whatever else the Arenas lab could think of.
