---
title: "Arenas lab name checker usage"
author: "Oliver Tam and Julie Lowndes"
date: "6/26/2020"
output: html_document
---

## Introduction

This document is to keep track of our workflow using `name_checker.R`,
a script to check that files conform to the Arenas lab convention: 

We expect 8 sections, each separated by an underscore `_`:

1. Experimental name and researcher initials
2. Experiment date and experiment number
3. Condition and replicate number
4. Immunohistochemistry date
5. Dye, antibodies or transcript
6. Image capture date
7. Microscope type
8. Lens, zoom and image number

The first 3 sections would be used (in a future script) to sort the
file into subfolders.

Example:

`NES-SAI2d15-CS_200514-01_Vehicle-1_200517_DAPI+goPitx3-dk488+rbAldh1-dk555+moTh-dk647_200519_CF_10Xz1-1.tiff`

Where:

- Experiment name & initial: `NES-SAI2d15-CS`. experiment name is NES-SAI2d15, CS is Carmen Salto
- Experiment date and number: `200514-01`. 14 May 2020, number 1
- Condition & replicate: `Vehicle-1`. Vehicle condition, replicate 1
- Immunohistochemistry date: `200517`: 17 May 2020
- Dye/antibodies/transcript:
  `DAPI+goPITX3-dk488+rbALDH1-dk555+moTH-dk647`. Each element
  separated by `+`, while primary and secondary antibody pairs are
  separated by `-`.
- Image capture date: `200520`: 20 May 2020
- Microscope type: `CF`. Confocal microscope 
- Lens, zoom & image number: `10X-z2-3`. 10x lens, 2 zoom, number 3.

## Code setup

Download or clone the GitHub folder: <https://github.com/olivertam/NDCN_hackathon>. Move it somewhere you do your analyses (we can discuss strategic options later).

Open the GitHub folder from RStudio File > Open Project... or in Finder/Windows Explorer by double-clicking the `NDCN_hackathon.Rproj` file.

Navigate to the "arenas-02" folder.

We will work from this `name_checker_usage.Rmd` document today. It is
an R Markdown document, which combines simple text and R code. You can
use the Table of Contents at the bottom of the file to navigate. And
please add your own notes!

We will load one library, `here`, which reduces potential filepath issues. But if you have trouble with the `here` package, don't worry, we'll provide an option without it throughout this demo.

```{r setup}
## checks if the `here` package is installed, if not, installs 
if(! requireNamespace("here",quietly=TRUE)){
  install.packages("here")
}

## load libraries
library(here) # load `here` package
```

## How to use the `name_checker` function

First we have to tell R we want to access code in the `name_checker.R` file, which we do with the `source()` function. This R script has a function called `name_checker()` that we'll use today.

Run the following code block if you have the `here` package:

```{r source_with_here}
source(here("arenas-02", "code", "name_checker.R"))
```

Here is the non-`here` code option:

```{r source_no_here, eval=FALSE}
source(file.path("arenas-02", "code", "name_checker.R"))
```

This code gives you a list of parameters that the `name_checker()` function would take:

```{r usage}
args(name_checker)
```

The following are some examples on how we can call this function:

```{r examples, eval=FALSE}
results <- name_checker("my_folder")

results <- name_checker("my_folder", print2screen=FALSE)
```

Above, the user provides a folder (`my_folder`) containing files with names
that should be checked. The `name_checker()` function finds all the image files (ending in .czi .tif or .tiff) and checks their names. 

The first command saves the results to a variable (called `results`), and also prints it to the screen. This is what we will be using for most of
this demo. 

The second command saves the results to a variable (`results`), but
doesn't print it to screen. This is useful if you don't want to see
the results on screen (especially if there are 100s of files to
process).

## Oliver's demo

There is a file in the `files_test` folder that is named according to
the Arenas lab convention. We can run the function to check if it is named correctly.

```{r oliver_test}
results <- name_checker(here("arenas-02",'files_test'))
```

```{r oliver_test_no_here, eval=FALSE}
results <- name_checker(file.path("arenas-02",'files_test'))
```

The function indicates that the file is named according to the proposed
nomenclature.

## Carmen's test run

In another folder, we had Carmen and Shanzheng from the Arenas lab try
to name a few files according to the proposed nomenclature. We have
also added some file names that might not perfectly fit. Let's see if
the function is able to check them.

```{r carmen_test}
source(here("arenas-02","code", "name_checker.R"))

results <- name_checker(here("arenas-02","files_Arenas"))
```

```{r carmen_test_no_here}
source(file.path("arenas-02","code", "name_checker.R"))

results <- name_checker(file.path("arenas-02","files_Arenas"))
```

Let's discuss the results.

### Questions?

Does anyone have any questions? Have you been following along with the code?

## Features discussion

We've created this code through discussions with the Arenas lab. Their lab discussed about what their current personal naming conventions are, and what nice lab conventions could be. Deciding on team norms and writing code to address them a very iterative process, and will continue after the hackathon.

These are some updates/iterations that we can discuss. We can also look at the code of what we would do to change it: 

- **File types**. The function is designed to only look at certain file types. Can we add to it?
- **Underscores**. The 8 sections are separated by underscores. Do we still like this? What if we wanted to change it to dashes? 
- **FirstCaps**. Currently gene names are in ALLCAPS. But what if we wanted them in FirstCaps? 
- **Dyes and antibodies**. Combining dyes and antibodies gets quite long and complicated. What are other options?
- **Verbose mode**. We can also use the `verbose` argument to give more information to the user.
- **Checking multiple folders**. The function currently operates on just one folder you provide. But we could have the code look through multiple folders.

## Understanding and tweaking the function

Before we take a look at the code, this is to remind you of the
nomenclature:

- 8 sections, each separated by an underscore `_`:

1. Experimental name and researcher initials
2. Experiment date and experiment number
3. Condition and replicate number
4. Immunohistochemistry date
5. Dye, antibodies or transcript
6. Image capture date
7. Microscope type
8. Lens, zoom and image number

And to remind you how we run this function:

```{r reminder, eval=FALSE}
results <- name_checker("my_folder")
```

Currently, the function is given a folder, and searches for all the
files within. It then takes each of the file names, and checks that it
matches the nomenclature based on the rules above. If it is incorrect,
it will try to highlight (and sometimes suggest) sources of error.

While it might seem that the function is designed specifically for the Arenas lab, there are a few features that could be re-used for other code.

### File types

One feature of this function is to only look at files that are either
images (.tif) or those generated by the microscope (.lsm or
.czi). This is because we don't want to be checking files where
we are not expecting them to use this naming system.
The code that selects those particular file types is on [line 102 of the `name_checker.R` script](https://github.com/olivertam/NDCN_hackathon/blob/master/arenas-02/code/name_checker.R#L102)

```{r find_files, eval=FALSE}
files <- setdiff(list.files(folder, pattern="\\.czi$ | \\.tif$ | \\.tiff$ | \\.lsm$",full.names = TRUE), list.dirs(recursive = FALSE))
```

If you look at the code, you can see that there is a series of
"patterns" that it is looking for. 
Let's take one example:

```
\\.czi$
```

This is a [regular
expression](https://en.wikipedia.org/wiki/Regular_expression), which
is a way to search for a pattern in text. Even if you don't know
anything about regular expressions, you might be guess what pattern
it's looking for based on what we expect the code to do.

What it is looking for in the example above are items (in this case, files) that end in ".czi". 

- The "." is represented as `\\.` (it's because both back-slashes (`\`) and dots (`.`) have special meaning in regular expression, so you have to use a weird combination of them to make the computer understand exactly what you want)
- The "czi" file extension is represented as `czi`
- The `$` indicates the end of the value, so nothing should come
  after it

#### Questions

How many different file extensions is it looking for? (hint: the `|`
operator means "or")

If you wanted to add more file extensions/type (let's say ".txt"), how
would you change the code?

### Underscores

Once the function has the file name, it tries to split the name into chunks by
treating the underscore as a separator. So if you have `a_b_c`, the
code on [line
126](https://github.com/olivertam/NDCN_hackathon/blob/master/arenas-02/code/name_checker.R#L126)
will break it up into `a`, `b`, and `c`

```{r breaking_into_chunks, eval=FALSE}
file_info = unlist(strsplit(name, "_", fixed=TRUE))
```

#### Question

If you decided that you want to split that information using commas as
a separator, how would you change the code?
E.g. You want `a,b,c` to become `a`, `b`, and `c`

*Note*: This is actually quite similar (in concept) to how many programs (e.g. Excel) turn a csv
(comma separated values) file into a table of "cells", though they
tend to have additional checks for the "unusual" cases.

### FirstCaps

One "feature" that is in the code is to turn all gene symbols into
upper-case. This makes sense for protein nomenclature, but sometimes you might prefer
having it with only first letter capitalized (especially if you're
working more with non-human genes).

How would you change that?

#### Combining code

We have another piece of code that would capitalize the first letter
of any word:

```{r source_firstCaps}
source(here("arenas-02","code","firstCaps.R"))
```

```{r source_firstCaps_no_here, eval=FALSE}
source(file.path("arenas-02","code","firstCaps.R"))
```

```{r firstCaps_demo}
args(firstCaps)

firstCaps("NURR1")
firstCaps("ras")
```

How would we add this into our function?

Let's head back into [`name_checker.R`](https://github.com/olivertam/NDCN_hackathon/blob/master/arenas-02/code/name_checker.R)

If we head to [line
92](https://github.com/olivertam/NDCN_hackathon/blob/master/arenas-02/code/name_checker.R#L92),
we find a function called `toupper()` there. 
Based on the name, what do you think the function does?

```{r toupper_test}
args(toupper)

toupper("NURR1")
toupper("ras")
```

*Note*: You also see this function on [line 88](https://github.com/olivertam/NDCN_hackathon/blob/master/arenas-02/code/name_checker.R#L88)

So now we find where the code is making things uppercase. How would we
change it to first caps?

First, we want to copy the `firstCaps()` code into the `name_check.R` script (which we
will save as a new copy called `name_checker_firstCaps.R`). I like to
paste it at the top.

Then we will change the code to use the `firstCaps()` function instead
of `toupper()`, and then save the file.

*Note*: make sure to change both line 88 and 92 (see below)

```{r firstCaps_change, eval=FALSE}
target = substr(antibodies[i],3,nchar(antibodies[i]))
if(grepl("^p[A-Z]",target)){
    ## Possible phosphorylated protein
    ## Converts the expected gene name to all uppercase,
    ##  but keeping the lower case "p" to indicate phosphorylated
    target = paste0("p",firstCaps(substr(target,2,nchar(target))))
}
else{
    ## This converts the expected gene name to all uppercase
    target = firstCaps(target)
}
```

Let's see what the code does now:

```{r firstCaps_nameChecker_source}
source(here("arenas-02","code","name_checker_firstCaps.R"))
```

```{r firstCaps_nameChecker_no_here, eval=FALSE}
source(file.path("arenas-02","code","name_checker_firstCaps.R"))
```

```{r firstCaps_nameChecker_demo}
results <- name_checker(file.path("arenas-02","files_test"))
```

Now you can see that the function is now suggesting the user to rename
the file so that only the first letter is capitalized for gene symbols.

We hope that this might demonstrate how you can tweak and/or combine
pieces of code from different sources, and make a new piece of code
that would do what you want to do.

### Dyes and antibodies

The Arenas lab work with a lot images using multiple antibodies and
dyes, and so they wish to include a lot of that information in their
file name. 

Here's one example:
`DAPI+goPITX3-dk488+rbALDH1-dk555+moTH-dk647`

This indicates that the image is stained with DAPI, and treated with
the following combinations of primary & secondary antbodies:

- goat anti-PITX3 with donkey fluorophore-488
- rabbit anti-ALDH1 with donkey fluorphore-555
- mouse anti-TH with donkey fluorphore-647

Currently, the functions expects the different dyes and antibodies to
be separated by pluses (`+`), and for antibody labeling to be in the
form: 
(primary antibody)-(secondary antibody)
We are also providing a two-letter code for the host species
generating the antibody.

That is a lot of information contained in this section of the file
name, and thus it's probably something that the function should check
a little more carefully.

Currently, the code handles this by splitting the long name into
smaller chunks (using the `+` as the separator) (see [line 209](https://github.com/olivertam/NDCN_hackathon/blob/master/arenas-02/code/name_checker.R#L209)

So `DAPI+goPITX3-dk488+rbALDH1-dk555+moTH-dk647`
becomes
`DAPI`, `goPITX3-dk488`, `rbALDH1-dk555` ,`moTH-dk647`

It then tries to find a dash (`-`) in each chunk, and if a dash is
found, it's treated as an antibody label.

Once it decides it's an antibody, it then splits the primary from the
secondary: `goPITX3-dk488` becomes `goPITX3` and `dk488`.
The function then extracts the first two letter to get the host species ID (e.g. `go`),and treats the rest as the target (gene symbol or fluorphore). As we discussed above, it capitalizes the gene symbol if required.

Unfortunately, the function currently can't confirm that the
information is correct, but merely tries to extract it. Considering
how many pieces of information is contained here, it might be worth trying to check it more closely.

#### Questions

- The function assumes that having a `-` means an antibody pair. What happens
  if this is an in-situ hybridization with a gene that has a dash in its name? 
    - Should we use another symbol (like `=` or `#`)?
    - The file names would start looking quite ridiculous
- What happens if you forget to add the host species to the antibody?
     - The gene symbol or flurophore might now be misread
- Are there other ways to store this information?

### Verbose mode

Even though the function says that it's named correctly, sometimes you
might want to check what it's actually finding. The code also has a way to do this by invoking the `verbose` option:

```{r verbose}
results  <- name_checker(here("arenas-02","files_test","Good_file"),verbose=TRUE)
```

```{r verbose_no_here}
results  <- name_checker(file.path("arenas-02","files_test","Good_file"),verbose=TRUE)
```

From the results, you can see that it identified the following:
1. Destination folder
2. Experimental condition
3. Replicate number (if present)
4. Dye, antibody or transcripts being detected
5. Date of image capture
6. Type of microscope used
7. Lens used
8. Zoom level
9. Picture number

So, in addition to checking the file name is correct, the function
could also extract information that could be used in the future to:

1. Provide a destination folder for a "file moving" code
2. Generate a  file to store the image information (metadata)
3. Find all the files that go into the same folder, and generate a
   randomized name for each file corresponding to the IHC experiment
   - This would allow for "blinded" image analysis.

### Checking multiple folders

This function can be used as part of a larger workflow to check
multiple folders.
The following R code lists all the directory in the current folder
(which will include all three folders with image files), and will run
the function on each of those folders.
Furthermore, it will also save the results into a file called
`name_checking_log.txt` that you can look at once all the folders have been checked

```{r multiFolder, eval=FALSE}
all_dirs  <-  list.dirs(path=file.path("arenas-02"))
for(i in 1:length(all_dirs)){
  logfile  <-  c(paste("Current folder is:",all_dirs[i]),"")
  logfile  <-  c(logfile,name_checker(all_dirs[i]),print2screen=FALSE)
}

write.table(logfile,file.path("arenas-02","name_checking_log.txt"),sep="\n",quote=F,row.names=F,col.names=F)
```





