---
title: "Arenas lab file renaming usage"
author: "Julie Lowndes & Oliver Tam"
date: "5/25/2020"
output: html_document
---

## Introduction

This document is to keep track of our workflow using `batch_rename_files.R`, a script to rename files within the working directory folder. 

## Setup

Download the GitHub folder. Move it somewhere you do your analyses (we can discuss good options later).  

Look inside: it has a .Rproj file; double-click that .Rproj file to open RStudio to this working directory. We will be working today in RStudio.

We will work from this `arenas_lab_usage.Rmd` document today. It is an R Markdown document, which combines simple text and R code. You can use the Table of Contents at the bottom of the file to navigate. And please add your own notes!

## Look through the script

Let's open the `batch_rename_files.r` script and get a sense of what it does. 

The `batch_rename_files.r` script has 1 main function:

`batch_rename_files`. It requires the user to provide inputs (in quotes!) for these variables:

- `microscope`, e.g. "BF"
- `antibody`, e.g. "Thrb488dkDAPI"
- `picture_date`, e.g. "20200522"
- `lens_magnification`, e.g. "20x2_0" 

This function will: 

1. It will require user input to confirm how you set variables. We'll do this from the R Console. 
2. It will replace illegal characters: `[/\\?<>\\:*|\":]`
3. It will create new filename prefix and log information 
4. It will identify files to rename
5. Finally, it will write information in the log file

## Using the script

Let's try it back from this document. We need to "call" this function, which we can do from another place where we document our workflow.

Start off by RESTARTING R. File menu > Session > Restart R. 

(yes we just opened R with RStudio but it's a great habit to get into!)

```{r setup}

## libraries ---
## `here` package helps manage filepaths so that this will work more easily on different computers

## checks if the `here` package is installed and usable
## installs `here` package if not installed
if(! require(here)){
    install.packages("here")
}

library(here) # load `here` package

## source the script so you can use it
source(file.path("hackathon-arenas-01","code", "batch_rename_files.r"))

```

Now let's use the script. We will rename the example image files in this repo.

We will need to change the **working directory** to where the image files we want to process are.

```{r usage, eval = FALSE}

## change your working directory to where the files are with `setwd()`
#setwd(here("rename_files_examples", "test_files"))

## rename files 
batch_rename_files(microscope = "BF",
                   antibody = "Thrb488dkDAPI", 
                   picture_date = "20200522",
                   lens_magnification = "20x2_0")

```

This will return files called, for example, `CF-Thrb488dkDAPI_20200522_20X2_0-01.jpg`, which has the format `microscope–antibody_YYYYMMDD_magnification–#.jpg`

What happens if you type `n` to the prompts?

## Testing on Carmen's computer

Let's try this again on Carmen's computer and discuss!

Now that it works on Carmen's computer, we can start talking about how it could be improved.


## Improvements 

What are ideas we'd like to improve?

### Running R script on multiple folders

Can we run this script on a bunch of directories?

We could provide several inputs for microscopes, antibodies, picture_dates, and lens_magnifications, and also several inputs for the directories filepaths. Here is what this could look like: 

```{r multi_directory, eval=FALSE}

## Store current directory to get back to it at the end
current_directory <- getwd()

## This can be set 10 when you run the function (i.e. in this RMarkdown file) or  2) provided as an input file
directory_list <- c("dirA", "dirB", "dirC")           # Option 1
directory_list <- read.csv("list_of_directories.csv") # Option 2

## Parameters for renaming files can be provided as four separate vectors
microscopes <- c("BF","CF","EM")
antibodies <- c("Thrb488dkDAPI","Exosc3488nucHoechst","mtDNAgold")
picture_dates <- c("20200529","20200101","20191130")
lens_magnifications <- c("20x2_0","40x","TEM")

## loop through directory_list ----
for(index in 1:length(directory_list)){
  if(dir.exists(directory_list[index])){
    setwd(directory_list[index])
    batch_rename_files(microscope = microscopes[index],
                       antibody = antibodies[index],
                       picture_date = picture_dates[index],
                       lens_magnification = lens_magnifications[index])
  }
  else{
    warning(paste("This directory", directory_list[index], "does not exist. Skipping"))
  }
}
setwd(current_directory)

```

We could also add these checks before the for loop above to make sure that our input is correct:

```{r multi_directory_checks, eval=FALSE}
# Check that the parameter lists and directory lists are the same ----
if(length(directory_list) != length(microscopes)){
  stop(paste("There are", length(directory_list), "directories given, but only", length(microscopes), "microscope information provided"), call.=FALSE)
}
if(length(directory_list) != length(antibodies)){
  stop(paste("There are", length(directory_list), "directories given, but only", length(antibodies), "antibody information provided"), call.=FALSE)
}
if(length(directory_list) != length(picture_dates)){
  stop(paste("There are", length(directory_list), "directories given, but only", length(picture_dates), "picture dates provided"), call.=FALSE)
}
if(length(directory_list) != length(microscopes)){
  stop(paste("There are", length(directory_list), "directories given, but only", length(lens_magnifications), "lens magnification information provided"), call.=FALSE)
}

```

## R script functionality

- Do we want to rename the script and function?
- Do we want to rename the "Converted" folder?
- What if we wanted to reorder the name so that the date came just before the counter, e.g.:  `CF-Thrb488dkDAPI_20X2_0_20200522-01.jpg`

We would focus on the `prefix` variable in `batch_rename_files.R`: 

```{r prefix, eval=FALSE}
 ## Create new filename prefix and log information ----
  prefix <- paste(microscope, antibody, picture_date, lens_magnification, sep="_")
  prefix <- sanitize(prefix)
```


## Improving documentation and naming

- Let's add more comments to the function and to this .Rmd document so that you feel comfortable using it. 
- try using `----` in your code chunk comments so that you can navigate at the bottom of the file editor


### Adding function documentation with ROxygen

1. Open `batch_rename_files.R`
1. Put your cursor in the function somewhere (e.g. below line 6)
1. Menu bar > Code > Insert Roxygen skeleton
1. Fill in the parameters with descriptions

### Helper messages

Should we have it print a message when completed, identifying the log file?


## Discussion

What are your questions about this workflow? 

### Working directory

Where the script is vs. where the files are vs. where your working directory is
- where should this script "live" on Carmen's computer?
  - maybe a folder called "analysis" from her home directory

### Naming things

**Dates** and [naming things](https://speakerdeck.com/jennybc/how-to-name-files): recommend YYYYMMDD not YYMMDD. (even better: YYYY-MM-DD)

Recommend not using "." because can be confused with file extension

## More practice 

### What is wrong here: 

What happens?

```{r error, eval = FALSE}

rename_files(microscope = "BF",
             antibody, = "Thrb488dkDAPI", 
             picture_date = "20200522",
             lens_magnification = "20x2_0")


```


