---
title: "Arenas lab name checker usage"
author: "Oliver Tam and Julie Lowndes"
date: "6/26/2020"
output: html_document
---

## Introduction

This document is to keep track of our workflow using `name_checker.R`, a script to check that files conform to the Arenas lab convention: 



## Setup

Download the GitHub folder. Move it somewhere you do your analyses (we can discuss good options later).  

Look inside: it has a .Rproj file; double-click that .Rproj file to open RStudio to this working directory. We will be working today in RStudio.

We will work from this `arenas_lab_usage.Rmd` document today. It is an R Markdown document, which combines simple text and R code. You can use the Table of Contents at the bottom of the file to navigate. And please add your own notes!


## Testing using name_checker.R

```{r setup}
## checks if the `here` package is installed, if not, installs 
if(! require(here)){
  install.packages("here")
}

## load libraries
library(here) # load `here` package

## source the script so you can use it
source(here("hackathon-arenas-02", "code", "name_checker.R"))
```


## check and rename files

```{r}
logfile <- check_file_names(here('hackathon-arenas-02/files'))
```


### Test run

```{r test-run}
all_dirs = list.dirs()
logfile = vector()
for(i in 1:length(all_dirs)){
  cat(paste("Current folder is:",all_dirs[i]), "\n\n")
  logfile = c(paste("Current folder is:",all_dirs[i]),"")
  logfile = c(logfile,check_file_names(all_dirs[i]))
}
write.table(logfile,"name_checking_log.txt",sep="\n",quote=F,row.names=F,col.names=F)

```





